---
title: "C2 - Assignment 2."
author: "Son Nam Nguyen"
date: "11/27/2021"
output:
  prettydoc::html_pretty:
  toc: true
  theme: architect
  highlight: github
---

```{r setup, include=FALSE}

if (!require(pacman)){
  install.packages("pacman")
}

pacman::p_load(rvest, httr, tidyverse, jsonlite, data.table, tidyr, stringr, patchwork, kableExtra)

```

## Access API

I have decided to scrape the TendersGuru public API, which contains procurement data
of the countries situated in the Pannonian Basin along with Spain. Please, find the
documentation on the API [here](https://tenders.guru/hu/api).

```{r, message=F, warning=F}
#send GET request
res <- httr::GET(url = 'https://tenders.guru/api/hu/tenders')

#retrive JSON response and flatten out the object
t <- fromJSON(content(res, 'text'), flatten = T)

#convert it to dataframe
df <- data.frame(t$data)

df %>%
  select(-c("description", "awarded")) %>% #drop long text column for clarity
  head(5) %>% 
  kbl() %>%
  kable_styling("striped") %>%
  scroll_box(width = "100%")

```

</br>

## Processing

What we can observe from the original data transformed from the JSON object,
is that there are nested dataframes in in the `awarded` column of the original dataset.
Beyond that, there are also further nested dataframes inside the dataframes of the `awarded` column
called as `suppliers` containing the slug of tender suppliers.

```{r}

#lets extract the list of dataframes from each row and bind them together with rbindlist
awarded <-
  rbindlist(lapply(t$data$awarded, function(x){
    data.frame(x, stringsAsFactors = F)
  }), fill=T)

#we can unnest the resulted dataframe by supplier ids
awarded <- unnest(awarded, suppliers)

#Before we move onto the data visualization,
#we have to make the necessary datatype transformations. 

#fixing datatypes of the purchaser
df <- df %>% mutate(date = as.POSIXct(date), #convert char to date
                    awarded_value_m = as.numeric(awarded_value)/1000000, #change metric to millions
                    deadline_length_days = as.integer(deadline_length_days), #cast days to integer
                    category = as.factor(category)) #cast category to factor

#fixing datatypes of the supplier
awarded <- awarded %>% mutate(value_m = as.numeric(value)/1000000) #change metric to millions

```

At the end of the day, the flattened out awarded dataframe looks as follows:

```{r}
awarded %>%
  head(5) %>%
  kbl() %>%
  kable_styling("striped") %>%
  scroll_box(width = "100%")

```

</br>

## DataViz

I decided to plot four charts altogether:

1. Share of categories between different phases of the procurement
2. Deadline days distribution
3. Historical movement of project values
4. Biggest suppliers

### Phase distribution across categories

Majority of the projects across categories are all in the contracting and performance phase,
and **only around 5% of them are ready for evaluation**.

```{r pressure, echo=FALSE, fig.align='center'}

ggplot(data = df, aes(fill=phase_en, x=category)) + 
  geom_bar(position='fill') +
  labs(title = 'Phases by Categories',
       x = '',
       y = 'Share') +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_viridis_d(name = 'Categories') +
  theme_bw()+
  theme(legend.position = "top")
```

### Distribution of deadline days

Looking at deadline length distribution, I decided to plot a histogram which indicates
that the shape is right-skewed and the **average due days are around 21 days for each procurement**.
The shaded rectangle presents deadline values which are the extreme values of the distribution.

```{r, fig.align='center', warning=F}

ggplot(df, aes(deadline_length_days)) +
  geom_histogram(colour="firebrick4", fill="firebrick1", bins=35) +
  geom_vline(aes(xintercept=mean(deadline_length_days, na.rm=T)),
             color="firebrick3", linetype="dashed", size=1, alpha = 0.5) +
  geom_text(aes(mean(deadline_length_days, na.rm=T), 15, label=paste0("Mean: ",round(mean(deadline_length_days, na.rm=T), 2))), hjust = -0.5) +
  geom_rect( inherit.aes=FALSE, aes(xmin=50, xmax=max(deadline_length_days, na.rm=T),ymin=0,
                                    ymax=20),  color="transparent", fill="lightsalmon", alpha=0.02) +
  labs(title = 'Distribution of deadline length',
       x = 'Days due',
       y = 'Frequency') +
  theme_bw()
  
```

### Movement of project values

We can recognize occasional spikes when looking at the change of awarded tender values.
**The highest spike was around August and September of 2021 moving beyond the value of 1 billion HUF**.
Furthermore, data points indicate that tenders have been more frequently advertised as we move
along the time axis. 

```{r, fig.align='center'}
ggplot(df, aes(date, awarded_value_m, color = "firebrick3"))+
  geom_line() +
  geom_point() +
  labs(title = "Change in Awarded Procurements",
       x = "Date",
       y = "Value (in M HUF)") +
  theme_bw() +
  theme(legend.position = "none")
```

### Biggest suppliers on the market

Lastly, I looked at the organizations with the highest totals in terms of tender values.
**The most successful procurement company was DRYVIT PROFI construction company** with a
whooping total beyond 1 billion HUF in value.

```{r, fig.align='center'}

awarded %>% 
  group_by(name) %>%
  summarise(total=sum(value_m)) %>%
  arrange(-total) %>% 
  slice(c(1:5)) %>% 
     ggplot( aes(reorder(stringr::str_wrap(name, 20), total), total)) +
        geom_col(colour="firebrick4", fill="firebrick1") +
        coord_flip() +
        labs(title = "Top 5 Suppliers",
             x = "",
             y = "Value (in M HUF)") +
        theme_bw()  

```
