
---
title: "Web scraping week 1"
output:
  md_document:
    toc: true
    variant: markdown_github
  output_file: README.md
---

```{md, echo=F, include=F, eval=F}

<details>
  <summary>Summary Goes Here</summary>
  ...this is hidden, collapsable content...
</details>
  
  
```

# Functions

```{r}
welcome <- function(name) {
  return(paste0('Warm welcome ', name))
}

welcome(name = 'Ceu')

```

**Task**<br>
Create a function called `multi` which will take argument `a` and `b` and return with the `a * b - b`
```{r, echo=F}
multi <- function(a, b) {
  return((a*b)-b)
}
```

```{r}
multi(2,5)
```

**Task**<br>
Create a function named `bmi_calc` which will have two inputs `weight`(kg) and `height`(cm) and calculate the BMI index `weight/height(meter)^2`. Check your result [here](https://www.nhlbi.nih.gov/health/educational/lose_wt/BMI/bmicalc.htm)

```{r, echo=F}
bmi_calc <- function(weight, height ) {
  return(weight /((height/100)^2) )
}

```


```{r}
bmi_calc(weight = 75, height = 175)
```




# Loops

```{r}
for (i in 1:10) {
  print(i)
}

j <-1
while (j<10) {
  t_url <- paste0('https://baseurl.com/page', j)
  print(t_url)
  j <- j+1
}


```

**Task**<br>
Reproduce the following printing with a for loop. (Use builtin vector called `letters`)
```{r, echo=FALSE}
for (i in 1:length(letters)) {
  print(paste0('The ', i, '. letter is ', letters[i]))
}


```



Create a function which will return the square of the input number.

```{r, echo=FALSE}
my_square <- function(x) {
  return(x^2)
}

```

```{r}
my_square(11)
```

# Sapply, lapply

Sapply and lapply takes two element, first a list and second a function<br>


Sapply returns with a vector
```{r}
sapply(1:10, my_square)
```


Lapply returns with a list
```{r}
lapply(1:10, my_square)
```


```{r}
lapply(1:10, function(x){x^3})
```

unlist will flat the list into a vector
```{r}
unlist(lapply(1:10, function(x){x^3}))

```


# HTML structure

* [elements](https://developer.mozilla.org/en-US/docs/Web/HTML/Element)
* [div](https://developer.mozilla.org/en-US/docs/Web/HTML/Element/div)
* [id](https://developer.mozilla.org/en-US/docs/Web/HTML/Global_attributes/id)

Load any site, press F12<br>
In the console you can select elements with `document.querySelector('')`. <br>
You can get the text of any div, element or id with `innerText` attribute `document.querySelector('h1').innerText`. <br>
You can also overwrite the text of the selected element `document.querySelector('h1').innerText=='hello'` It will change the first h1 text to hello. <br>

Try out with elements, id, and div.



**Task**<br>
Change two element on [this](https://www.guinnessworldrecords.com/world-records/most-lasso-texas-skips-in-one-minute) site

* who should contain your name
* where should contain your country
* take a screenshoot


`div` use `.` <br>
`id` use `#` <br>
`tag` use just the name of the tag.




# Rvest

Install and load the rvest library
```{r}
#install.packages("rvest")
library(rvest)
# load the index.html page
t <- read_html('index.html')

# tags
t %>% html_node('h1') %>% html_text()
t %>% html_nodes('h1') %>% html_text()

t %>% html_node('.demo-class') %>% html_text()

t %>% html_nodes('#select-with-id') %>% html_text()

```


```{r}
t <- read_html('https://www.boats.com/boats/prestige/420-8040261/')
t %>% html_nodes('#specifications .oem-page__title') %>% html_text()
t %>% html_nodes('.description-list__term') %>% html_text()
t %>% html_nodes('.description-list__description') %>% html_text()

```

